Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-05-05 doc/RMD/mlg-mcg.Rmd
fix issue with mlgs

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  as_data_frame() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  group_by(MCG) %>%
  summarize(MLGs = sum(Freq > 0), 
            Samples = sum(Freq), 
            Evenness = diversity_stats(Freq)["E.5"], 
            data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
  arrange(desc(MLGs))
mlgs <- mcgmlg %>%
  group_by(MLG) %>%
  summarize(MCGs = sum(Freq > 0), 
            Samples = sum(Freq), 
            Evenness = diversity_stats(Freq)["E.5"], 
            data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
  arrange(desc(Samples), desc(MCGs))
mcgs
mlgs

@@

@@

```{r barplots, fig.width = 10, fig.height = 4}
mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

@@

V(g)$size <- sqrt(osize)/10
set.seed(2017-05-05)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-04)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
```

>>>>

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  mutate(MLG = as.character(MLG)) %>%
  mutate(MCG = as.character(MCG)) %>%
  as_data_frame() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  arrange(MCG) %>%
  group_by(MCG) %>%
  mutate(MLGs = sum(Freq > 0), 
         Samples = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data = list(data_frame(MLG = MLG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(MLGs)) %>%
  ungroup() %>%
  distinct(MCG, MLGs, Samples, Evenness, data)
mlgs <- mcgmlg %>%
  arrange(MLG) %>%
  group_by(MLG) %>%
  mutate(MCGs     = sum(Freq > 0), 
         Samples  = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data     = list(data_frame(MCG = MCG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(Samples), desc(MCGs)) %>%
  ungroup() %>%
  distinct(MLG, MCGs, Samples, Evenness, data)
mcgs
mlgs

@@

These tables are a good start, but we want to have a publication-ready table:

```{r production-tables, results = 'asis'}
fct2int <- function(x) as.integer(as.character(x))

psex_from_graph <- function(g, MLG){
  weights <- setNames(E(g)$weight, E(g)$label)
  weights <- split(weights, names(weights))[MLG]
  1 -vapply(weights, unique, numeric(1))
}

mlg_table <- mlgs %>%
  slice(1:5L) %>%
  mutate(MLG = as.character(MLG)) %>%
  mutate(Psex = psex_from_graph(graph11loc$total, MLG)) %>%
  unnest() %>%
  group_by(MLG, Samples, Psex) %>%
  summarize(MCG = paste(MCG, collapse = ", ")) %>%
  rename(N = Samples) %>%
  arrange(desc(N))

kable(mlg_table, caption = "Top 5 MLGs", format.args = list(digits = 3))

mcg_table <- mcgs %>%
  slice(1:5L) %>%
  mutate(MCG = as.character(MCG)) %>%
  unnest() %>%
  group_by(MCG, Samples, Evenness) %>%
  summarize(MLG = list(strwrap(paste(MLG, collapse = ", "), 28))) %>%
  rename(N = Samples) %>%
  arrange(desc(N)) %>% 
  unnest() %>%
  ungroup() %>%
  mutate(Evenness = ifelse(!duplicated(MCG), format(Evenness, digits = 3), "")) %>%
  mutate(N = ifelse(!duplicated(MCG), N, "")) %>%
  mutate(MCG = ifelse(!duplicated(MCG), MCG, ""))
kable(mcg_table, caption = "Top 5 MCGs")
```

@@

```{r barplots, fig.width = 10, fig.height = 4}
mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

@@

V(g)$size <- sqrt(osize)/10
set.seed(2017-05-03)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-08)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
```
